---
title: "Proportion of schools with no sanitation service"
subtitle: "Exploring the data from Unicef through Total Obervation value since 2000"
author: "Edward Arby"
date: 2024/04/14
format: html
theme: quartz
backgroundcolor: lightgreen
execute: 
  echo: false
  warning: false
  message: false
---



```{r}
#| label: setup
#| include: false

# libraries
library(tidyverse)
library(plotly)

#data
unicef_indicator_1 <- read_csv("unicef_indicator_1.csv") 
```

The data on the proportion of schools with no sanitation service in various countries. One story that stands out is the situation in Mali, where 19.1% of schools have no sanitation service. This is a significant issue as access to proper *sanitation facilities* is crucial for maintaining hygiene and health, especially in schools where children are more susceptible to diseases.

```{r}
#| fig-cap: "World map showing the propotion of schools with no sanitation services"
map_world <- map_data("world")
map_data_join <-full_join(unicef_indicator_1,map_world, by = c("country" = "region"))

ggplot(map_data_join) +
  aes(x = long, y = lat, group = group, fill = obs_value) +  
  geom_polygon()
```

*Mali* has the highest proportion of schools with no sanitation service among the countries listed in the spreadsheet.This is in contrast to countries like *Malaysia, Mauritius, and Monaco*, where there are no schools without sanitation services.

```{r}
#| fig-cap: "Time-series sheet showing the propotion of schools with no sanitation services since 2000"
timeseries_plot_1 <-unicef_indicator_1 %>%
   
  ggplot() +
  aes (obs_value,time_period, color = country) +
  geom_line()

ggplotly(timeseries_plot_1)
```

Other countries with high proportions of schools with no sanitation service include *Madagascar (28.01%), Mauritania (60.43%), and Niger (64.39%)*.

```{r}
#| fig-cap: "Scatter plot sheet showing the propotion of schools with no sanitation services since 2000"
scatter_plot_1 <- unicef_indicator_1 %>%
ggplot() +
  aes(obs_value, time_period, color = country) +
  geom_point()

ggplotly(scatter_plot_1)
```


```{r}
#| fig-cap: "Bar chart representing the propotion of schools with no sanitation services since 2000"
bar_chart_1 <- unicef_indicator_1 %>%
  ggplot() +
  aes(alpha_3_code, obs_value, fill = country) +
  geom_col()

ggplotly(bar_chart_1)  
```

The situation in Mali and other countries with high proportions of schools without sanitation services underscores the need for urgent action to address this issue. Providing access to *proper sanitation facilities* in schools can help *prevent the spread of diseases* and *promote better health outcomes for children*.
